---
title: "ggplot2: Scale x log"
author: "Vebash Naidoo"
date: "06/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(scales)
library(flair)
```

```{r eda1, cache=TRUE}
fishing <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')

fishing %>% 
  summarise(max_gt = max(grand_total,
                 na.rm = TRUE),
            min_gt = min(grand_total,
                 na.rm = TRUE))

fishing %>% 
  count(grand_total, sort = TRUE)


fishing %>% 
  count(grand_total)

# quite skewed because of many 0's and
# large values in the grand_total
fishing %>% 
  ggplot(aes(grand_total)) +
  geom_histogram()
```

As we see there is much data all around `0`. To see it a bit clearer we can use a log transform on the x-axis data. 

<span style="color: #5196b4;background-color:#f4e4e7">`scale_x_log10()`</span> puts the x-axis on a log scale.

Note: There is also similarly a <span style="color: #5196b4;background-color:#f4e4e7">`scale_y_log10()`</span>.

```{r scale-x-1, include=FALSE}
# scale the x axis to using log values
# to see the distribution better
# we see many 0's and then a few other numbers
fishing %>% 
  ggplot(aes(grand_total + 1)) +
  geom_histogram() +
  scale_x_log10()
```


```{r, echo = FALSE}
decorate_chunk('scale-x-1') %>% 
  flair('scale_x_log10()', 
        background = "#f4e4e7", 
        color = "#5196b4")
```


We can also set the <span style="color: #5196b4;background-color:#f4e4e7">breaks</span> ourselves.

```{r scale-x-1-1, include=FALSE}
# scale the x axis to using log values
# to see the distribution better
# also set the breaks
fishing %>% 
  ggplot(aes(grand_total + 1)) +
  geom_histogram() +
  scale_x_log10(breaks = c(0, 10, 100, 1000, 10000, 20000, 50000))
```

```{r, echo = FALSE}
decorate_chunk('scale-x-1-1') %>% 
  flair('breaks = c(0, 10, 100, 1000, 10000, 20000, 50000)', 
        background = "#f4e4e7", 
        color = "#5196b4")
```


## Example 2

Here's another example taken from Rebecca Barter's [blog post on ggplot2](https://www.rebeccabarter.com/blog/2017-11-17-ggplot2_tutorial/).

```{r scale-x-2}
gapminder <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")

gapminder_2007 <- gapminder %>% filter(year == 2007)

ggplot(gapminder_2007, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point(alpha = 0.5)
```

Again interesting data is near 0 and hard to see. Let's scale the x axis again. This time we use <span style="color: #5196b4;background-color:#f4e4e7">`label = comma`</span> to make our axis labels have a <span style="color: #5196b4;background-color:#f4e4e7">`,`</span>: `1,000`, `2,000`, `3,000` etc.

```{r scale-x-3, include = FALSE}
ggplot(gapminder_2007, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point(alpha = 0.5) +
  scale_x_log10(label = comma)
```

```{r, echo = FALSE}
decorate_chunk('scale-x-3') %>% 
  flair('label = comma', 
        background = "#f4e4e7", 
        color = "#5196b4")
```


